---
title: "Berekley SDA HTML Page Parser"
author: "Kieran healy"
date: "October 13th, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



```{r libraries}
library(tidyverse)
library(rvest)
```

# What this is

The Berkeley SDA archive (<http://sda.berkeley.edu>) has a number of public datasets such as the GSS, together with HTML versions of their codebooks. For the [GSSR package](http://kjhealy.github.io/gssr), I scraped the codebook pages in order to get them into a tibble that could be bundled with the package. It seems that the SDA has almost identical codebooks for the other surveys it carries, so this code could be adapted for use with them. There will be some differences---e.g. the GSS has a "Text of this Question" field along with marginal summaries of the variable for each question in the survey, while the ANES seems to lack this "Text of Question" part. But they're very close and the HTML/CSS document structure is the same. 

Here's the code for the GSS documentation. The repo includes the raw HTML files, so the code should run.

# Initial data scraping of codebook and saving to local storage

## Scrape the GSS codebook from SDA

This next code chunk shows how we got the codebook data, but it is not evaluated here (we set `echo = FALSE`), because we only need to do it once.

```{r get-remote-pages, eval = FALSE, echo = TRUE}

## Generate vector of doc page urls
urls <- paste0("https://sda.berkeley.edu/D3/GSS18/Doc/", 
               "hcbk", sprintf('%0.4d', 1:261), ".htm")


## Grab the codebook pages one at a time
doc_pages <- urls %>% 
  map(~ {
    message(glue::glue("* parsing: {.x}"))
    Sys.sleep(5) # try to be polite
    safely(read_html)(.x)
  })

```

## Save the scraped webpages locally

There's a gotcha with objects like `doc_pages`: they cannot be straightforwardly saved to R's native data format with `save()`. The XML files are stored with external pointers to their content and cannot be "serialized" in a way that saves their content properly. If you try, when you `load()` the saved object you will get complaints about missing pointers. So instead, we'll unspool our list and save each fellow's page individually. Then if we want to rerun this analysis without crawling everything again, we will load them in from our local saved versions using `read_html()`.

Again, this code chunk is shown but not run, as we only do it once. 

```{r localsave, eval = FALSE, echo = TRUE}

## Get a list containing every codebook webpage, 
## Drop the safely() error codes from the initial scrape (after we've checked them), 
## and also drop any NULL entries
page_list <- pluck(doc_pages, "result") %>% 
  compact()

## Make a vector of clean file names of the form "raw/001.htm"
## One for every page we grabbed. Same order as the page_list.
## We use sprintf to get numbers of the form 001, 002, 003 etc.
fnames <-paste0("raw/", 
                sprintf('%0.4d', 1:length(doc_pages)),
                ".htm") 

## Walk the elements of the page list and the file names to 
## save each HTML file under is respective local file name
walk2(page_list, fnames, ~ write_xml(.x, file = .y))

```

# Read in the pages from the local directory

Using the local data we've saved, we read in a list of all the web pages. Our goal is to get them into a tractable format (a tibble or data frame). From there we can write some functions to, e.g., query the codebook directly from the console, or alterantively produce the codebook in a format suitable for integrating into the R help system via a package.

```{r localparse}

## The names of all the files we just created
local_urls <- fs::dir_ls("raw/")

## Read all the pages back in, from local storage 
doc_pages <- local_urls %>% 
  map(~ {
    safely(read_html)(.x)
  })

## Are there any errors?
doc_pages %>% pluck("error") %>% 
  flatten_dfr()

## quick look at first five items in the list
summary(doc_pages)[1:5,]

## Quick look inside the first record
doc_pages[[1]]

```

# Parse the pages

Next, we parse every webpage to extract a row for every variable. There are multiple variables per page.

## Functions

```{r functions}

## Page of variables to list of variables and their info, 
parse_page <- function(x){
  html_nodes(x, ".dflt") %>%
    map(~ html_nodes(.x, ".noborder")) %>%
    map(~ html_table(.x))
}

## Length of each list element
## Standard GSS Qs will have 4 elements
## Ids recodes and other things will have 3
get_lengths <- function(x){
  map(x, length)
}

get_names <- function(x){
  map(x, names)
}

## Variable short names and descriptions
get_var_ids <- function(x){
  x %>% map_dfr(1) %>%
    select(id = X1, description = X3) %>%
    as_tibble()
}


## Question Text
get_text <- function(x, y){
  if(y[[1]] == 3) {
    return(NA_character_)
  } else {
    stringr::str_trim(x[[2]])
  }
}

## Question Marginals
get_marginals <- function(x, y){
  if(y[[1]] == 3) {
    tmp <- x[[2]]
  } else {
    tmp <- x[[3]]
  }
  
  if(ncol(tmp) == 2) {
    as_tibble(tmp) %>%
      select(cases = X1, range = X2)
  } else {
    tmp <- as_tibble(tmp[, colSums(is.na(tmp)) != nrow(tmp)]) %>%
      janitor::clean_names()
    tmp$value <- as.character(tmp$value)
    tmp
  }
}

## Add an id column
add_id <- function(x, y){
  x %>% add_column(id = y)
}

## Question Properties
get_props <- function(x, y){
  if(y[[1]] == 3) {
    tmp <- x[[3]]
    colnames(tmp) <- c("property", "value")
    tmp <- as_tibble(tmp)
    tmp$property <- stringr::str_remove(tmp$property, ":")
    tmp
  } else {
    tmp <- x[[4]]
    colnames(tmp) <- c("property", "value")
    tmp <- as_tibble(tmp)
    tmp$property <- stringr::str_remove(tmp$property, ":")
    tmp
  }
}

## Take the functions above and process a page to a tibble of cleaned records

process_page <- function(x){
  page <- parse_page(x)
  q_vars <- get_var_ids(page)
  lens <- get_lengths(page)
  keys <- q_vars$id
  
  q_text <- map2_chr(page, lens, ~ get_text(.x, .y))
  q_text <- stringr::str_trim(q_text)
  q_text <- stringr::str_remove_all(q_text, "\n")
  q_text <- tibble(id = keys, q_text = q_text)
  q_text <- q_text %>%
    mutate(q_text = replace_na(q_text, "None"))
  q_marginals <- map2(page, lens, ~ get_marginals(.x, .y)) %>%
    set_names(keys) 
  q_marginals <- map2(q_marginals, keys, ~ add_id(.x, .y))
  
  q_props <- map2(page, lens, ~ get_props(.x, .y)) %>%
    set_names(keys) 
  q_props <- map2(q_props, keys, ~ add_id(.x, .y))
  
  q_tbl <- q_vars %>% 
    add_column(properties = q_props) %>% 
    add_column(marginals = q_marginals) %>%
    left_join(q_text) %>%
    rename(text = q_text)
  
  q_tbl

  }

```

## Make the tibble

Parse the GSS variables into a tibble, with list columns for the marginals and the variable properties.

```{r create-tibble-be-patient}
gss_doc <-  doc_pages %>% 
  pluck("result") %>% # Get just the webpages
  compact() %>%
  map(process_page) %>%
  bind_rows()
```

## Look at the outcome

```{r outputs}

gss_doc
gss_doc$id <- tolower(gss_doc$id)

```

## 

```{r outputs2}
gss_doc %>% filter(id == "race") %>% 
  select(text)

gss_doc %>% filter(id == "race") %>% 
  select(marginals) %>% 
  unnest(cols = c(marginals))
```

```{r outputs3}
gss_doc %>% filter(id == "sex") %>% 
  select(text)

gss_doc %>% filter(id == "sex") %>% 
  select(marginals) %>% 
  unnest(cols = c(marginals))
```

```{r outputs4}
gss_doc %>% filter(id == "fefam") %>% 
  select(text)

gss_doc %>% filter(id == "fefam") %>% 
  select(properties) %>% 
  unnest(cols = c(properties))
```

# Save the data object as efficiently as we can

Shown here but not run

```{r save-object, eval = FALSE}
save(gss_doc, file = "data/gss_doc.rda", 
     compress = "xz") 
# tools::resaveRdaFiles("data")
# tools::checkRdaFiles("data")
```
